Crop diversification in a game against nature

### 2020-01-22 risk-percption.Rmd
remove extra plots

Given these payoffs, what's the point of indifference between the two crop types?
```{r, echo = FALSE}
indiff <- tibble(risk = (barley_wet - wheat_wet) / (wheat_dry - wheat_wet - barley_dry + barley_wet),
                 utility = risk * wheat_dry + (1 - risk) * wheat_wet)

tibble(risk = seq(0, 1, .1),
      barley_exp = risk * barley_dry + (1 - risk) * barley_wet,
      wheat_exp = risk * wheat_dry + (1 - risk) * wheat_wet) %>%
  gather(type, utility, barley_exp:wheat_exp) %>%
  ggplot(aes(risk, utility)) +
  geom_hline(yintercept = indiff$utility, linetype = 2)+
  geom_vline(xintercept = indiff$risk, linetype = 2) +
  geom_line(aes(color = type), size = 1.2) +
  geom_point(data = indiff, color = 'black', size = 3) +
  geom_text(data = indiff, label = '(0.3, 1.1)', nudge_x = 0.1, nudge_y = 0.1) +
  theme_classic() +
  ggtitle('Point of indifference between wheat and barley', 'Given varying drought risk')
```

@@

@@

tibble(wheat_prop = seq(0, 1, .1),
       dry_exp = wheat_prop * wheat_dry + (1 - wheat_prop) * barley_dry,
       normal_exp = wheat_prop * wheat_normal + (1 - wheat_prop) * barley_normal,
      wet_exp = wheat_prop * wheat_wet + (1 - wheat_prop) * barley_wet) %>%
  gather(type, utility, dry_exp:wet_exp) %>%
  ggplot(aes(wheat_prop, utility)) +
 # geom_hline(yintercept = nash_eq$utility, linetype = 2)+
#  geom_vline(xintercept = nash_eq$wheat_prop, linetype = 2) +
  geom_line(aes(color = type), size = 1.2) +
#  geom_point(data = nash_eq, color = 'black', size = 3) +
 # geom_text(data = nash_eq, 
#            label = paste0('(', nash_eq$wheat_prop, ', ', nash_eq$utility, ')'),
#            nudge_x = 0.1, nudge_y = 0.05) +
  theme_classic()
```

```{r}
tibble(wheat_prop = seq(0, 1, .1),
       dry_exp = wheat_prop * wheat_dry + (1 - wheat_prop) * barley_dry,
       normal_exp = wheat_prop * wheat_normal + (1 - wheat_prop) * barley_normal,
      wet_exp = wheat_prop * wheat_wet + (1 - wheat_prop) * barley_wet,
      yield_exp = dry_exp + normal_exp + wet_exp, 
      lower_limit = pmin(dry_exp, normal_exp,wet_exp)) %>%
  ggplot(aes(yield_exp, lower_limit)) +
 # geom_hline(yintercept = nash_eq$utility, linetype = 2)+
#  geom_vline(xintercept = nash_eq$wheat_prop, linetype = 2) +
  geom_line(aes(color = type), size = 1.2) +
#  geom_point(data = nash_eq, color = 'black', size = 3) +
 # geom_text(data = nash_eq, 
#            label = paste0('(', nash_eq$wheat_prop, ', ', nash_eq$utility, ')'),
#            nudge_x = 0.1, nudge_y = 0.05) +
  theme_classic()
```

@@

```{r}
test_dat <- expand_grid(wheat_prop = (0:10)/10,
       risk = seq(.1, .9, 0.1)) %>%
  mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
         util_dry = risk * (wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop)),
         util_wheat = wheat_prop * (wheat_wet * (1 - risk) + wheat_dry *  risk),
         util_barley = (1 - wheat_prop) * (barley_wet* (1 - risk) + barley_dry * risk),
         payoff_exp = util_dry + util_wet,
         payoff_max = pmax(util_wheat, util_barley))
```

```{r}
nash_eq2 <- tibble(risk = seq(0.1, 0.9, 0.1),
                   wheat_prop = pmax(round(((1 - risk) * barley_wet - risk * barley_dry) / (risk * wheat_dry + (1 - risk) * barley_wet - ((1 - risk) * wheat_wet + risk * barley_dry)), 2), 0),
                  utility = round(risk * wheat_dry * wheat_prop + risk * barley_dry * (1 - wheat_prop), 2))

test_dat %>%
  select(-(util_wheat:util_barley)) %>%
  gather(type, utility, c(util_dry, util_wet)) %>%
    ggplot(aes(wheat_prop, utility)) + 
  geom_line(aes(color = type)) + 
  facet_wrap(~risk) + 
  geom_hline(data = nash_eq2, aes(yintercept = utility), alpha = .3, linetype = 2) + 
  geom_vline(data = nash_eq2, aes(xintercept = wheat_prop), alpha = 0.3,linetype = 2) +
  geom_point(data = nash_eq2) +
  theme_minimal()
# hmmm, on second thought this might not make any sense, what's actually informative here is tha maximum value, which is the same as saying whether it is above or below the indifference threshold. but beyond that the idea of the nash equilibrium at different risk levels doesn't make sense, becuase then the nash equilibrium is, under the hood, adding another risk assessment to nature's moves.
```

```{r}
test_dat %>%
  select(-(util_dry:util_wet)) %>%
  gather(type, util, util_wheat:util_barley) %>%
  ggplot(aes(wheat_prop, util, color = type)) + 
  geom_line() + 
  facet_wrap(~risk) + 
 # geom_vline(xintercept = nash_eq$wheat_prop) +
  theme_minimal()
```

*Here we see that the mixed strategy that maximizes expected utility is always a pure strategy.*
```{r}
test_dat %>%
  gather(type, utility, payoff_exp:payoff_max) %>%
  ggplot(aes(wheat_prop, utility)) + 
  geom_line(aes(color = type)) +
  facet_wrap(~risk)
```

@@

sim_past <- brick('~/Downloads/b40.mh6ka.1deg.003.cam2.h0.PRECT.080101-131612.nc') %>% 
  raster::extract(konya, df = TRUE) %>%
  select(-ID) %>%
  gather(time, rainfall) %>%
  mutate(rainfall = rainfall * 2.628e+9,
         year = rep(1:516, each = 12), # add hardcode dates instead of using ncdf time because of date shift with averaging period
         month = rep(1:12, times = 516)) %>%
  select(-time) %>%
  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
   filter(!(month %in% c(5:9))) %>%
  group_by(water_year) %>%
  summarise(precip = sum(rainfall))
```

```{r}
obs <- read_table('~/Downloads/pa17244.dat', 
           comment = "#", 
           col_names = c('year', 1:12), 
           na = c('-999.9', '-888.8')) %>%
  gather(month, precip, `1`:`12`) %>%
  group_by(month) %>%
  mutate(precip = if_else(is.na(precip), mean(precip, na.rm = TRUE), precip)) %>%
  ungroup %>%
  mutate(month = as.numeric(month),
         water_year = if_else(month %in% 10:12, year + 1, year)) %>% 
    filter(between(water_year, 1932, 2005)) %>%
 filter(!(month %in% c(5:9))) %>%
  group_by(water_year) %>%
  summarise(precip = sum(precip))
```

>>>>

Given these payoffs, what's the point of indifference between the two crop types?
```{r, echo = FALSE}
indiff <- tibble(risk = (barley_wet - wheat_wet) / (wheat_dry - wheat_wet - barley_dry + barley_wet),
                 yield = risk * wheat_dry + (1 - risk) * wheat_wet)
indiff_label <- paste0('(', round(indiff$risk, 2), ', ', round(indiff$yield, 2), ')')

tibble(risk = seq(0, 1, .1),
      barley_exp = risk * barley_dry + (1 - risk) * barley_wet,
      wheat_exp = risk * wheat_dry + (1 - risk) * wheat_wet) %>%
  gather(type, yield, barley_exp:wheat_exp) %>%
  ggplot(aes(risk, yield)) +
  geom_hline(yintercept = indiff$yield, linetype = 2)+
  geom_vline(xintercept = indiff$risk, linetype = 2) +
  geom_line(aes(color = type), size = 1.2) +
  geom_point(data = indiff, color = 'black', size = 3) +
  geom_text(data = indiff, label = indiff_label, nudge_x = 0.1, nudge_y = 0.1) +
  theme_classic() +
  ggtitle('Point of indifference between wheat and barley', 'Given varying drought risk')
```

@@

```{r}
qplot(x = c(300, 400, 500), y = c(0, 1.5, 2), geom = 'line') +
  geom_line(aes(y = c(0.8, 1, 0.5)))
```

@@

tibble(wheat_prop = seq(0, 1, .1),
       dry_exp = wheat_prop * wheat_dry + (1 - wheat_prop) * barley_dry,
       normal_exp = wheat_prop * wheat_normal + (1 - wheat_prop) * barley_normal,
      wet_exp = wheat_prop * wheat_wet + (1 - wheat_prop) * barley_wet) %>%
  gather(type, utility, dry_exp:wet_exp) %>%
  ggplot(aes(wheat_prop, utility)) +
  geom_hline(yintercept = nash_eq$utility, linetype = 2)+
  geom_vline(xintercept = nash_eq$wheat_prop, linetype = 2) +
  geom_line(aes(color = type), size = 1.2) +
  geom_point(data = nash_eq, color = 'black', size = 3) +
  geom_text(data = nash_eq, 
            label = paste0('(', nash_eq$wheat_prop, ', ', nash_eq$utility, ')'),
            nudge_x = 0.1, nudge_y = 0.05) +
  theme_classic()
```

```{r}
tibble(wheat_prop = seq(0, 1, .1),
       dry_exp = wheat_prop * wheat_dry + (1 - wheat_prop) * barley_dry,
       normal_exp = wheat_prop * wheat_normal + (1 - wheat_prop) * barley_normal,
      wet_exp = wheat_prop * wheat_wet + (1 - wheat_prop) * barley_wet,
      yield_exp = dry_exp + normal_exp + wet_exp, 
      lower_limit = pmin(dry_exp, normal_exp,wet_exp)) %>%
  ggplot(aes(yield_exp, lower_limit)) +
 # geom_hline(yintercept = nash_eq$utility, linetype = 2)+
#  geom_vline(xintercept = nash_eq$wheat_prop, linetype = 2) +
  geom_line( size = 1.2) +
#  geom_point(data = nash_eq, color = 'black', size = 3) +
 # geom_text(data = nash_eq, 
#            label = paste0('(', nash_eq$wheat_prop, ', ', nash_eq$utility, ')'),
#            nudge_x = 0.1, nudge_y = 0.05) +
  theme_classic()
```

@@

*Here we see that the mixed strategy that maximizes expected utility is always a pure strategy.*

@@

sim_past <- brick('~/Downloads/b40.mh6ka.1deg.003.cam2.h0.PRECT.080101-131612.nc') %>% 
  raster::extract(konya, df = TRUE) %>%
  select(-ID) %>%
  gather(time, rainfall) %>%
  mutate(rainfall = rainfall * 2.628e+9,
         year = rep(1:516, each = 12), # add hardcode dates instead of using ncdf time because of date shift with averaging period
         month = rep(1:12, times = 516)) %>%
  select(-time) %>%
  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
   filter(!(month %in% c(6:9))) %>%
  group_by(water_year) %>%
  summarise(precip = sum(rainfall))
```

```{r}
obs <- read_table('~/Downloads/pa17244.dat', 
           comment = "#", 
           col_names = c('year', 1:12), 
           na = c('-999.9', '-888.8')) %>%
  gather(month, precip, `1`:`12`) %>%
  group_by(month) %>%
  mutate(precip = if_else(is.na(precip), mean(precip, na.rm = TRUE), precip)) %>%
  ungroup %>%
  mutate(month = as.numeric(month),
         water_year = if_else(month %in% 10:12, year + 1, year)) %>% 
    filter(between(water_year, 1932, 2005)) %>%
 filter(!(month %in% c(6:9))) %>%
  group_by(water_year) %>%
  summarise(precip = sum(precip))
```
